---
title: "P8105_HW02_LF2649"
author: "Lin_Feng_LF2649"
date: "September 24, 2020"
output: 
      github_document:
        toc : true
---

```{r setup}
library(tidyverse)
library(readxl)
```

#Problem 1

## Read the Mr.Traswheel dataset

```{r}
trashwheel_df = read_xlsx(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel", 
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
```

## Read precipatation data for 2018 and 2017

```{r}
precip_2018=
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation", 
    skip = 1 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)



precip_2017=
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation", 
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Combine annual precipitation: 

```{r}
month_df = tibble(
    month = 1:12, 
    month_name = month.name
)

precip_df = bind_rows(precip_2018, precip_2017)

precip_df = left_join(precip_df, month_df, by = "month")
```

The dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. 
As trash enters the inner harboe, the trashwheel collects that trash, and stores it in a dumster. 
The dataset contains information on year, month, and trash collected, include some specific kinds of trash. 
There are a lot of `r nrow(trashwheel_df)` rows in pur final dataset. Additional datasheets include month and year. 
In this dataset: 

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


#Problem 2

##Import, clean, and tidy the NYC Transit data. Convert "entry". 

```{r import_clean}
 nyc_trans = 
  read_csv("./NYC_Transit.csv", 
           col_types = cols(
             Route8 = col_character(),
             Route9 = col_character(),
             Route10 = col_character(),
             Route11 = col_character()
           )) %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  pivot_longer(
    route1:route11, 
    names_to = "route_name", 
    names_prefix = "route_number", 
    values_to = "exit") %>% 
  
  mutate(entry = recode(entry, YES = "TRUE", NO = "FALSE")) 

```

##Description of the dataset 
This dataset contains information about each entrance and exit for each subway station in NYC. Data cleaning steps include formatting variable names into proper style, selecting needed variables, tidying the route names and number, and convert the character "entry" into logical variable. After the cleaning an dtidying, there are total `r nrow(nyc_trans)` rows and `r ncol(nyc_trans)` columns in the dataset. It is a tidy dataset with `r ncol(nyc_trans)`
variables: `r names(nyc_trans)`.

##Distinct statione, ADA, and wending
```{r}
a = distinct(nyc_trans, line, station_name) %>% 
    count()

b = filter(nyc_trans, ada == TRUE) %>%
    distinct(line, station_name) %>%
    count()

c = filter(nyc_trans, entry == TRUE, vending == "NO") %>%
    distinct(line, station_name) %>%
    count()
```

* There are `r distinct(nyc_trans, line, station_name) %>% count()` stations in NYC
* `r filter(nyc_trans, ada == TRUE) %>% distinct(line, station_name) %>% count()` stations are ADA compliant among those all stations in NYC. 
* The propotion of station entrances / exits without vending allow entrance is `r c/a`.


#Problem 3

##First, deal with pols-month.csv

* clean the data in pols-month.csv. 
* Use separate() to break up the variable mon into integer variables year, month, and day; 
replace month number with month name; 
create a president variable taking values gop and dem, and remove prez_dem and prez_gop; 
and remove the day variable.
```{r pols_month}
pols_month = read_csv("./5308/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "m_d"), sep = 4) %>% 
  separate(m_d, c(month",  "day"), sep = 4) %>% 
  mutate(year = as.double(year),day = as.double(day))
  
  
```

